### 2015-05-10 swmpr_manu.tex
all acronyms defined using glossaries package, added glossary to beginning of text

Two approaches can be used to import \ac{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \ac{CDMO} web services. To do so, the IP address for the computer making the request must be registered by following instructions on the \ac{CDMO} website.  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \ac{CDMO}.

% table for retrieval functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} for individual functions or \\texttt{help.search(`retrieve', package = `SWMPr')} for all).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:retrieve",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?all\_params} for individual functions or \texttt{help.search(`retrieve', package = `SWMPr')} for all).\label{tab:retrieve}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{all\_params}&Retrieve records starting with the most recent at a given station, all parameters.  Wrapper to \texttt{exportAllParamsXMLNew} function on web services.\tabularnewline
\texttt{all\_params\_dtrng}&Retrieve records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \texttt{exportAllParamsDateRangeXMLNew}.\tabularnewline
\texttt{import\_local}&Import files from a local path.  The files must be in a specific format, such as those returned from the \ac{CDMO} using the zip downloads option.\tabularnewline
\texttt{single\_param}&Retrieve records for a single parameter starting with the most recent at a given station.  Wrapper to \texttt{exportSingleParamXMLNew} function on web services.\tabularnewline
\texttt{site\_codes}&Metadata for all stations, wrapper to \texttt{exportStationCodesXMLNew} function on web services.\tabularnewline
\texttt{site\_codes\_ind}&Metadata for all stations at a single site, wrapper  to \texttt{NERRFilterStationCodesXMLNew} function on web services.\tabularnewline
\hline
\end{tabular}\end{center}

@@

% attributes table
%latex.default(to_tab[, -1, drop = F], file = "", caption = "Attributes of a \\texttt{swmpr} object that describe characteristics of the data.",     rowlabel = "Attributes", colheads = c("Class", "Description"),     rowname = to_tab$Attributes, caption.loc = "top", insert.bottom = foot,     col.just = c("p{0.75in}", "p{3.25in}"), label = "tab:attributes",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Attributes of a \texttt{swmpr} object that describe characteristics of the data.\label{tab:attributes}} 
\begin{center}
\begin{tabular}{lp{0.75in}p{3.25in}}
\hline\hline
\multicolumn{1}{l}{Attributes}&\multicolumn{1}{c}{Class}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{names}&character&Column names of the entire data set, inherited from the \texttt{data.frame} object class\tabularnewline
\texttt{row.names}&integer&Row names of the data set, inherited from the \texttt{data.frame} object class\tabularnewline
\texttt{class}&character&Class of the data object indicating \texttt{swmpr} and \texttt{data.frame}\tabularnewline
\texttt{station}&character&Station identifier used by \ac{NERRS} as a string with 7 or 8 characters\tabularnewline
\texttt{parameters}&character&Character vector of column names for data parameters, e.g., \texttt{`do\_mgl'}\tabularnewline
\texttt{qaqc\_cols}&logical&Indicates if \ac{QAQC} columns are present in the raw data\tabularnewline
\texttt{date\_rng}&POSIXct&Start and end dates for the raw data\tabularnewline
\texttt{timezone}&character&Timezone of the station using the city/country format\textsuperscript{a}\tabularnewline
\texttt{stamp\_class}&character&Class of the \texttt{datetimestamp} column, usually POSIXct unless data have been aggregated\tabularnewline
\hline
\end{tabular}\end{center}

@@

The \texttt{qaqc} function is a simple screen to retain observations from the data with specified QAQC flags (see \url{http://cdmo.baruch.sc.edu/data/qaqc.cfm}). Each parameter in the imported \texttt{swmpr} object will have a corresponding QAQC column of the same name with the added prefix \texttt{f\_} (e.g., \texttt{do\_mgl}, \texttt{f\_do\_mgl}).  Values in the QAQC column range from -5 to 5 to indicate the \ac{QAQC} flag that was assigned by \ac{CDMO} during initial processing.  The \ac{QAQC} function is used to remove observations in the raw data with given flags, with the default option to retain only values with the \texttt{0} \ac{QAQC} flag (i.e., passed initial \ac{CDMO} checks).     Additionally, simple filters are used to remove obviously bad values, e.g., wind speed values less than zero or pH values greater than 12. Erroneous data entered as -99 are also removed. The function returns the original data with the QAQC columns removed and \texttt{NA} (not available) values for observations that do not meet the criteria specified in the function call.

% table for organize functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb} for individual functions or \\texttt{help.search(`organize', package = `SWMPr')} for all).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:organize",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?comb} for individual functions or \texttt{help.search(`organize', package = `SWMPr')} for all).\label{tab:organize}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{comb}&Combines \texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.\tabularnewline
\texttt{qaqc}&Remove \ac{QAQC} columns and remove data based on \ac{QAQC} flag values for a \texttt{swmpr} object.  Only applies if \ac{QAQC} columns are present. \tabularnewline
\texttt{qaqcchk}&View a summary of the number of observations in a \texttt{swmpr} object that are assigned to different \ac{QAQC} flags used by \ac{CDMO}.  The output can be used to inform further processing.\tabularnewline
\texttt{rem\_reps}&Remove replicate nutrient data that occur on the same day.  The default is to average replicates.\tabularnewline
\texttt{setstep}&Format data from a \texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \texttt{comb} and can also be used with individual stations.\tabularnewline
\texttt{subset}&Subset by dates and/or columns for a \texttt{swmpr} object.  This is a method passed to the generic \texttt{subset} function provided in the base package.\tabularnewline
\hline
\end{tabular}\end{center}

@@

Several additional functions are provided that do not fit the above categories (\Cref{tab:misc}).  These functions are generally used within the main functions but may be useful for more customized evaluation of \ac{SWMP} data.

% table for misc functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Miscellaneous functions available from the SWMPr package.  Most are used within the main functions above but may be useful for customized evaluations of \\ac{SWMP} data.  Full documentation for each function is in the help file (e.g., execute \\texttt{?calckl} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:misc",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Miscellaneous functions available from the SWMPr package.  Most are used within the main functions above but may be useful for customized evaluations of \ac{SWMP} data.  Full documentation for each function is in the help file (e.g., execute \texttt{?calckl} at the command line).\label{tab:misc}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{calckl}&Estimate the reaeration coefficient for air-sea gas exchange.  Used in the \texttt{ecometab} function.\tabularnewline
\texttt{metab\_day}&Identify the metabolic day for each approximate 24 period in an hourly time series.  Used in the \texttt{ecometab} function.\tabularnewline
\texttt{param\_names}&Returns column names as a list for the parameter types (nutrients, weather, or water quality).  Includes \ac{QAQC} columns with \texttt{f\_} prefix. Used in the data retrieval functions.\tabularnewline
\texttt{parser}&Parses HTML returned from \ac{CDMO} data requests.  Used in the retrieval functions.\tabularnewline
\texttt{swmpr}&Creates a \texttt{swmpr} object class.  Used in the data retrieval functions.\tabularnewline
\texttt{time\_vec}&Converts time vectors to \texttt{POSIXct} objects with the appropriate time zone for a site.  Used in the data retrieval functions.\tabularnewline
\hline
\end{tabular}\end{center}

@@

This section describes three examples using the SWMPr package to illustrate the improved ability to synthesize and evaluate multi-year time series of estuarine data. First, the open-water method for estimating metabolism was applied to nearly all co-located water quality and weather sites at each \ac{NERRS} reserve using all years of available data.  The results are provided primarily to illustrate ease of use of the functions and secondarily to provide an update on metabolism estimates using the most recent \ac{SWMP} data.  Caffrey \cite{Caffrey03} and Caffrey \cite{Caffrey04} applied the open-water method to estimate ecosystem metabolism using five years of water quality observations at two sites at each of the \ac{NERRS} reserves.  The air-sea gas exchange model also assumed a constant value for the reaeration coefficient, whereas the current \texttt{ecometab} function provides a more accurate estimate by including weather data in the calculation (see Caffrey et al. \cite{Caffrey14} for details).

Water quality and weather observations from January 1995 to December 2014 for all \ac{NERRS} sites were obtained through a bulk data request using the zip downloads feature of \ac{CDMO}.  All csv files for each station were imported into R using the \texttt{import\_local} function, processed using the \texttt{setstep} and \texttt{qaqc} functions, then saved locally as binary RData files (see \nameref{met_scripts}). This resulted in a single \texttt{swmpr} object for each parameter at each site.  All files were then uploaded to a remote server for online access.  An additional R script (see \nameref{met_scripts}) was executed that retrieved and processed water quality and weather data at each reserve to estimate metabolism.  Two water quality sites with the longest time series at each reserve were used.  Mean annual values at each site, organized by region, are shown in \Cref{fig:metab_plo}, whereas decadal comparisons are shown in \Cref{tab:metab_tab}.  All sites were generally net heterotrophic across the range of observations (i.e., sink of organic matter, in agreement with Caffrey \cite{Caffrey03}), although differences were observed in early (i.e., 1995-2004) as compared to recent (2005-2014) time periods. Overall, the results indicate that between-region and within-site differences in metabolism are apparent and varying by time period, such that a more comprehensive evaluation of factors that influence metabolic rates is needed.  More importantly, use of the data retrieval, synthesis, and analysis functions to create the results illustrates the utility provided by the SWMPr package.

@@

\caption[Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve]{Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve.  Values are daily integrated estimates as mean annual values averaged across all years with 95\% confidence intervals.  Two sites were chosen from each reserve that had the longest available time series. Sites were assigned to regions based on approximate geographic coordinates.}\label{fig:metab_plo}
\end{figure}

@@

\caption{Trends in metabolism for two sites at each of the \ac{NERRS} reserves.  Values are averages of mean annual estimates for each period of observation (1994-2004 and 2005-2014). Bold values indicate an increase from the first period, whereas italic values indicate a decrease. Sites were assigned to regions based on approximate geographic coordinates.\label{tab:metab_tab}} 

@@

The second and third examples are two interactive web applications \cite{Chang15} created using the SWMPr package that illustrate summaries and comparisons of \ac{SWMP} data.  The first web application evaluates trends in \ac{SWMP} data within and between sites using an interactive map (\cref{fig:swmp_comp}): \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}.  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-level trends are shown below the map with a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend, such that larger points with darker colors indicate a strong trend.  The trend direction is blue for decreasing and red for increasing with time.  The second application provides graphical summaries of water quality, weather, or nutrient station data at individual stations using the \texttt{plot\_summary} function: \href{https://beckmw.shinyapps.io/swmp_summary/}{https://beckmw.shinyapps.io/swmp\_summary}.  The output is identical to \cref{fig:summary_ex}. Drop down menus can be used to select the station, date range, and parameter for plotting. The data used for each application are similar to those used to estimate ecosystem metabolism described above and are available in the \nameref{supp_info}.  Both of the apps provide previously unavailable utilities to interact with \ac{SWMP} data using an intuitive online interface.  This format may be more appealing for individuals with limited experience using R, while simultaneously illustrating what is possible using functions from the package. The applications have been used extensively with over 300 hours noted in a single month.

\begin{figure}
\begin{center}
\includegraphics[width = 0\textwidth]{figure/swmp_comp.pdf}
\caption{Online application for comparing trends in \ac{SWMP} data parameters using an interactive map.  Link: \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}}
\label{fig:swmp_comp}
\end{center}
\end{figure}

@@

The ability of management and research programs to address critical environmental issues is highly dependent on the quality of data used to inform decision making.  Standardized monitoring programs have vastly improved the ability to evaluate factors that influence a range of conditions, leading to more comprehensive assessments of site-specific characteristics and more informed decisions to manage environmental resources.  The \acl{SWMP} has provided twenty years of continuous monitoring of environmental characteristics at over over 140 stations within the 28 estuaries of the \acl{NERRS}.  This monitoring network establishes a foundation  for more effective coastal management by providing standardized data to address spatiotemporal variation in natural and anthropogenic characteristics that influence environmental condition.  Although the data provided by \ac{SWMP} are unique among coastal observing systems and have been used in a variety of applications \cite{Bulthius95,Caffrey03,Sanger02,Wenner04,Dix08}, the capacity of \ac{NERRS} researchers and staff to more effectively evaluate \ac{SWMP} data could be greatly improved using the SWMPr package.

The SWMPr package provides several functions to retrieve, organize, and analyze \ac{SWMP} data to more effectively address common challenges working with large datasets.  The package is designed to augment, rather than replace, existing data retrieval programs by providing a bridge betwen the raw data and the analysis software through its numerous data retrieval functions (\Cref{tab:retrieve}).  Established \ac{QAQC} methods and data processing techniques are also enhanced with SWMPr by functions that filter observations for different \ac{QAQC} flags (\texttt{qaqc}) and subset by selected dates or variables (\texttt{subset}).  Additionally, cumbersome challenges comparing differents datasets are addressed by the \texttt{setstep} and \texttt{comb} functions that standardize time steps and combine the data.  Finally, the analysis functions provide numerous tools to implement common analyses for time series and more specific methods for water quality data.  In particular, the \texttt{ecometab} function can be used to estimate daily integrated rates of ecosystem metabolism using the open-water method \cite{Odum56,Caffrey14}.  The above analysis (see \nameref{swmp_apps}) provided a cursory update of metabolism estimates for each the \ac{NERRS} estuaries using recent data to evaluate trends over time.  Although further evaluation of the data are needed, particularly regarding assumptions of the open-water method and tidal effects, the results could be used in a more comprehensive evaluation of factors that influence estuary metabolism. Further development of the SWMPr package will consider modifying existing and including additional functions to more effectively integrate data analysis with the quality of information provided by \ac{SWMP} and \ac{NERRS}.

@@

\href{https://s3.amazonaws.com/swmpexdata/zip_ex.zip}{https://s3.amazonaws.com/swmpexdata/zip\_ex.zip} A compressed data folder is provided to show the format provided by the zip downloads feature of the \ac{CDMO}.  These data can be retrieved for use in R with the \texttt{import\_local} function.  The same data are included with the SWMPr package as binary RData files.

\subsection*{\ac{SWMP} data as R binary files}

\href{https://s3.amazonaws.com/swmpalldata/}{https://s3.amazonaws.com/swmpalldata/} All \ac{SWMP} data from 1995 through 2014, one file per site. See the help file for \texttt{import\_remote} for more information.

@@

I acknowledge the significant work of \ac{NERRS} researchers and staff that has allowed access to high-quality monitoring data.  Thanks to Dwayne Porter and Melissa Ide from \ac{CDMO} for maintaining the online database.  Thanks to Marie Bundy and Nikki Dix for giving me the opportunity to share this package with the broader \ac{NERRS} community.  Thanks to Todd O'Brien for the inspiration for the online widgets. Thanks to Mike Murrell and Jim Hagy III for assistance with documentation and implementation of the ecosystem metabolism functions. The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S. Environmental Protection Agency.  The use of trade names or products does not constitute endorsement by the US Government.

>>>>

Two approaches can be used to import \gls{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \gls{CDMO} web services. To do so, the IP address for the computer making the request must be registered by following instructions on the \gls{CDMO} website.  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \gls{CDMO}.

% table for retrieval functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} for individual functions or \\texttt{help.search(`retrieve', package = `SWMPr')} for all).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:retrieve",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?all\_params} for individual functions or \texttt{help.search(`retrieve', package = `SWMPr')} for all).\label{tab:retrieve}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{all\_params}&Retrieve records starting with the most recent at a given station, all parameters.  Wrapper to \texttt{exportAllParamsXMLNew} function on web services.\tabularnewline
\texttt{all\_params\_dtrng}&Retrieve records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \texttt{exportAllParamsDateRangeXMLNew}.\tabularnewline
\texttt{import\_local}&Import files from a local path.  The files must be in a specific format, such as those returned from the \gls{CDMO} using the zip downloads option.\tabularnewline
\texttt{single\_param}&Retrieve records for a single parameter starting with the most recent at a given station.  Wrapper to \texttt{exportSingleParamXMLNew} function on web services.\tabularnewline
\texttt{site\_codes}&Metadata for all stations, wrapper to \texttt{exportStationCodesXMLNew} function on web services.\tabularnewline
\texttt{site\_codes\_ind}&Metadata for all stations at a single site, wrapper  to \texttt{NERRFilterStationCodesXMLNew} function on web services.\tabularnewline
\hline
\end{tabular}\end{center}

@@

% attributes table
%latex.default(to_tab[, -1, drop = F], file = "", caption = "Attributes of a \\texttt{swmpr} object that describe characteristics of the data.",     rowlabel = "Attributes", colheads = c("Class", "Description"),     rowname = to_tab$Attributes, caption.loc = "top", insert.bottom = foot,     col.just = c("p{0.75in}", "p{3.25in}"), label = "tab:attributes",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Attributes of a \texttt{swmpr} object that describe characteristics of the data.\label{tab:attributes}} 
\begin{center}
\begin{tabular}{lp{0.75in}p{3.25in}}
\hline\hline
\multicolumn{1}{l}{Attributes}&\multicolumn{1}{c}{Class}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{names}&character&Column names of the entire data set, inherited from the \texttt{data.frame} object class\tabularnewline
\texttt{row.names}&integer&Row names of the data set, inherited from the \texttt{data.frame} object class\tabularnewline
\texttt{class}&character&Class of the data object indicating \texttt{swmpr} and \texttt{data.frame}\tabularnewline
\texttt{station}&character&Station identifier used by \gls{NERRS} as a string with 7 or 8 characters\tabularnewline
\texttt{parameters}&character&Character vector of column names for data parameters, e.g., \texttt{`do\_mgl'}\tabularnewline
\texttt{qaqc\_cols}&logical&Indicates if \gls{QAQC} columns are present in the raw data\tabularnewline
\texttt{date\_rng}&POSIXct&Start and end dates for the raw data\tabularnewline
\texttt{timezone}&character&Timezone of the station using the city/country format\textsuperscript{a}\tabularnewline
\texttt{stamp\_class}&character&Class of the \texttt{datetimestamp} column, usually POSIXct unless data have been aggregated\tabularnewline
\hline
\end{tabular}\end{center}

@@

The \texttt{qaqc} function is a simple screen to retain observations from the data with specified QAQC flags (see \url{http://cdmo.baruch.sc.edu/data/qaqc.cfm}). Each parameter in the imported \texttt{swmpr} object will have a corresponding QAQC column of the same name with the added prefix \texttt{f\_} (e.g., \texttt{do\_mgl}, \texttt{f\_do\_mgl}).  Values in the QAQC column range from -5 to 5 to indicate the \gls{QAQC} flag that was assigned by \gls{CDMO} during initial processing.  The \gls{QAQC} function is used to remove observations in the raw data with given flags, with the default option to retain only values with the \texttt{0} \gls{QAQC} flag (i.e., passed initial \gls{CDMO} checks).     Additionally, simple filters are used to remove obviously bad values, e.g., wind speed values less than zero or pH values greater than 12. Erroneous data entered as -99 are also removed. The function returns the original data with the QAQC columns removed and \texttt{NA} (not available) values for observations that do not meet the criteria specified in the function call.

% table for organize functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb} for individual functions or \\texttt{help.search(`organize', package = `SWMPr')} for all).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:organize",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?comb} for individual functions or \texttt{help.search(`organize', package = `SWMPr')} for all).\label{tab:organize}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{comb}&Combines \texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.\tabularnewline
\texttt{qaqc}&Remove \gls{QAQC} columns and remove data based on \gls{QAQC} flag values for a \texttt{swmpr} object.  Only applies if \gls{QAQC} columns are present. \tabularnewline
\texttt{qaqcchk}&View a summary of the number of observations in a \texttt{swmpr} object that are assigned to different \gls{QAQC} flags used by \gls{CDMO}.  The output can be used to inform further processing.\tabularnewline
\texttt{rem\_reps}&Remove replicate nutrient data that occur on the same day.  The default is to average replicates.\tabularnewline
\texttt{setstep}&Format data from a \texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \texttt{comb} and can also be used with individual stations.\tabularnewline
\texttt{subset}&Subset by dates and/or columns for a \texttt{swmpr} object.  This is a method passed to the generic \texttt{subset} function provided in the base package.\tabularnewline
\hline
\end{tabular}\end{center}

@@

Several additional functions are provided that do not fit the above categories (\Cref{tab:misc}).  These functions are generally used within the main functions but may be useful for more customized evaluation of \gls{SWMP} data.

% table for misc functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Miscellaneous functions available from the SWMPr package.  Most are used within the main functions above but may be useful for customized evaluations of \\gls{SWMP} data.  Full documentation for each function is in the help file (e.g., execute \\texttt{?calckl} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:misc",     table.env = FALSE, where = "!tbp")%
\begin{table}[!tbp]
\caption{Miscellaneous functions available from the SWMPr package.  Most are used within the main functions above but may be useful for customized evaluations of \gls{SWMP} data.  Full documentation for each function is in the help file (e.g., execute \texttt{?calckl} at the command line).\label{tab:misc}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{calckl}&Estimate the reaeration coefficient for air-sea gas exchange.  Used in the \texttt{ecometab} function.\tabularnewline
\texttt{metab\_day}&Identify the metabolic day for each approximate 24 period in an hourly time series.  Used in the \texttt{ecometab} function.\tabularnewline
\texttt{param\_names}&Returns column names as a list for the parameter types (nutrients, weather, or water quality).  Includes \gls{QAQC} columns with \texttt{f\_} prefix. Used in the data retrieval functions.\tabularnewline
\texttt{parser}&Parses HTML returned from \gls{CDMO} data requests.  Used in the retrieval functions.\tabularnewline
\texttt{swmpr}&Creates a \texttt{swmpr} object class.  Used in the data retrieval functions.\tabularnewline
\texttt{time\_vec}&Converts time vectors to \texttt{POSIXct} objects with the appropriate time zone for a site.  Used in the data retrieval functions.\tabularnewline
\hline
\end{tabular}\end{center}

@@

This section describes three examples using the SWMPr package to illustrate the improved ability to synthesize and evaluate multi-year time series of estuarine data. First, the open-water method for estimating metabolism was applied to nearly all co-located water quality and weather sites at each \gls{NERRS} reserve using all years of available data.  The results are provided primarily to illustrate ease of use of the functions and secondarily to provide an update on metabolism estimates using the most recent \gls{SWMP} data.  Caffrey \cite{Caffrey03} and Caffrey \cite{Caffrey04} applied the open-water method to estimate ecosystem metabolism using five years of water quality observations at two sites at each of the \gls{NERRS} reserves.  The air-sea gas exchange model also assumed a constant value for the reaeration coefficient, whereas the current \texttt{ecometab} function provides a more accurate estimate by including weather data in the calculation (see Caffrey et al. \cite{Caffrey14} for details).

Water quality and weather observations from January 1995 to December 2014 for all \gls{NERRS} sites were obtained through a bulk data request using the zip downloads feature of \gls{CDMO}.  All csv files for each station were imported into R using the \texttt{import\_local} function, processed using the \texttt{setstep} and \texttt{qaqc} functions, then saved locally as binary RData files (see \nameref{met_scripts}). This resulted in a single \texttt{swmpr} object for each parameter at each site.  All files were then uploaded to a remote server for online access.  An additional R script (see \nameref{met_scripts}) was executed that retrieved and processed water quality and weather data at each reserve to estimate metabolism.  Two water quality sites with the longest time series at each reserve were used.  Mean annual values at each site, organized by region, are shown in \Cref{fig:metab_plo}, whereas decadal comparisons are shown in \Cref{tab:metab_tab}.  All sites were generally net heterotrophic across the range of observations (i.e., sink of organic matter, in agreement with Caffrey \cite{Caffrey03}), although differences were observed in early (i.e., 1995-2004) as compared to recent (2005-2014) time periods. Overall, the results indicate that between-region and within-site differences in metabolism are apparent and varying by time period, such that a more comprehensive evaluation of factors that influence metabolic rates is needed.  More importantly, use of the data retrieval, synthesis, and analysis functions to create the results illustrates the utility provided by the SWMPr package.

@@

\caption[Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \gls{NERRS} reserve]{Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \gls{NERRS} reserve.  Values are daily integrated estimates as mean annual values averaged across all years with 95\% confidence intervals.  Two sites were chosen from each reserve that had the longest available time series. Sites were assigned to regions based on approximate geographic coordinates.}\label{fig:metab_plo}
\end{figure}

@@

\caption{Trends in metabolism for two sites at each of the \gls{NERRS} reserves.  Values are averages of mean annual estimates for each period of observation (1994-2004 and 2005-2014). Bold values indicate an increase from the first period, whereas italic values indicate a decrease. Sites were assigned to regions based on approximate geographic coordinates.\label{tab:metab_tab}} 

@@

The second and third examples are two interactive web applications \cite{Chang15} created using the SWMPr package that illustrate summaries and comparisons of \gls{SWMP} data.  The first web application evaluates trends in \gls{SWMP} data within and between sites using an interactive map (\cref{fig:swmp_comp}): \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}.  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-level trends are shown below the map with a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend, such that larger points with darker colors indicate a strong trend.  The trend direction is blue for decreasing and red for increasing with time.  The second application provides graphical summaries of water quality, weather, or nutrient station data at individual stations using the \texttt{plot\_summary} function: \href{https://beckmw.shinyapps.io/swmp_summary/}{https://beckmw.shinyapps.io/swmp\_summary}.  The output is identical to \cref{fig:summary_ex}. Drop down menus can be used to select the station, date range, and parameter for plotting. The data used for each application are similar to those used to estimate ecosystem metabolism described above and are available in the \nameref{supp_info}.  Both of the apps provide previously unavailable utilities to interact with \gls{SWMP} data using an intuitive online interface.  This format may be more appealing for individuals with limited experience using R, while simultaneously illustrating what is possible using functions from the package. The applications have been used extensively with over 300 hours noted in a single month.

\begin{figure}
\begin{center}
\includegraphics[width = 0\textwidth]{figure/swmp_comp.pdf}
\caption{Online application for comparing trends in \gls{SWMP} data parameters using an interactive map.  Link: \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}}
\label{fig:swmp_comp}
\end{center}
\end{figure}

@@

The ability of management and research programs to address critical environmental issues is highly dependent on the quality of data used to inform decision making.  Standardized monitoring programs have vastly improved the ability to evaluate factors that influence a range of conditions, leading to more comprehensive assessments of site-specific characteristics and more informed decisions to manage environmental resources.  The \acrlong{SWMP} has provided twenty years of continuous monitoring of environmental characteristics at over over 140 stations within the 28 estuaries of the \acrlong{NERRS}.  This monitoring network establishes a foundation  for more effective coastal management by providing standardized data to address spatiotemporal variation in natural and anthropogenic characteristics that influence environmental condition.  Although the data provided by \gls{SWMP} are unique among coastal observing systems and have been used in a variety of applications \cite{Bulthius95,Caffrey03,Sanger02,Wenner04,Dix08}, the capacity of \gls{NERRS} researchers and staff to more effectively evaluate \gls{SWMP} data could be greatly improved using the SWMPr package.

The SWMPr package provides several functions to retrieve, organize, and analyze \gls{SWMP} data to more effectively address common challenges working with large datasets.  The package is designed to augment, rather than replace, existing data retrieval programs by providing a bridge betwen the raw data and the analysis software through its numerous data retrieval functions (\Cref{tab:retrieve}).  Established \gls{QAQC} methods and data processing techniques are also enhanced with SWMPr by functions that filter observations for different \gls{QAQC} flags (\texttt{qaqc}) and subset by selected dates or variables (\texttt{subset}).  Additionally, cumbersome challenges comparing differents datasets are addressed by the \texttt{setstep} and \texttt{comb} functions that standardize time steps and combine the data.  Finally, the analysis functions provide numerous tools to implement common analyses for time series and more specific methods for water quality data.  In particular, the \texttt{ecometab} function can be used to estimate daily integrated rates of ecosystem metabolism using the open-water method \cite{Odum56,Caffrey14}.  The above analysis (see \nameref{swmp_apps}) provided a cursory update of metabolism estimates for each the \gls{NERRS} estuaries using recent data to evaluate trends over time.  Although further evaluation of the data are needed, particularly regarding assumptions of the open-water method and tidal effects, the results could be used in a more comprehensive evaluation of factors that influence estuary metabolism. Further development of the SWMPr package will consider modifying existing and including additional functions to more effectively integrate data analysis with the quality of information provided by \gls{SWMP} and \gls{NERRS}.

@@

\href{https://s3.amazonaws.com/swmpexdata/zip_ex.zip}{https://s3.amazonaws.com/swmpexdata/zip\_ex.zip} A compressed data folder is provided to show the format provided by the zip downloads feature of the \gls{CDMO}.  These data can be retrieved for use in R with the \texttt{import\_local} function.  The same data are included with the SWMPr package as binary RData files.

\subsection*{\gls{SWMP} data as R binary files}

\href{https://s3.amazonaws.com/swmpalldata/}{https://s3.amazonaws.com/swmpalldata/} All \gls{SWMP} data from 1995 through 2014, one file per site. See the help file for \texttt{import\_remote} for more information.

@@

I acknowledge the significant work of \gls{NERRS} researchers and staff that has allowed access to high-quality monitoring data.  Thanks to Dwayne Porter and Melissa Ide from \gls{CDMO} for maintaining the online database.  Thanks to Marie Bundy and Nikki Dix for giving me the opportunity to share this package with the broader \gls{NERRS} community.  Thanks to Todd O'Brien for the inspiration for the online widgets. Thanks to Mike Murrell and Jim Hagy III for assistance with documentation and implementation of the ecosystem metabolism functions. The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S. Environmental Protection Agency.  The use of trade names or products does not constitute endorsement by the US Government.
